---
title: "Eating out"
author: "Kimberly Lemus MUnoz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data comes from the 1996 General Social Survey answer this question

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1996.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```



### Graphical Results
```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```

Based on the frequency table above, which shows the volume of people who eat out and people who do not eat out for each racial group, we can see that about 950 white people eat out while about 200 do not. Meanwhile, about 200 black people eat out while about 100 do not. For the other racial groups, about 100 people eat out while about 50 do not. Putting this into percentages,80% of white people eat out while 20% do not. Meanwhile black and others have about the same percentages with 70% of the people eat out while 30% do not.    

### Numerical Results

```{r}
table2 <- xtabs(~race + eatout,data=mydata)
rowPerc(table2)
colPerc(table2)
```
The numerical summary above shows the actual values of the percentages of people who eat out and do not eat out for each racial groups. The data above confirms the estimations from the graphical results, and it shows that for white people, 82.93% eat out and 17.17% do not. Similar to the observations made previously, black people and others have a similar ratio of about 70% of people who eat out while 31% of people who do not. Furthermore, the table above shows that white people make up the majority of people who claim that they eat out. Specifically, among the people answered yes to eating out, 82.94% were white people while 11.55% were black and 5.51% were others. 


###Inferential results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
Based on the Chi-squared test above, the p-value determined is a small number which means that there is a difference between the frequency of eating out of the the three racial groups. Therefore, we can reject the null hypothesis. To determine specifically which groups are different, another type of test will be conducted. 




```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(mydata$race,mydata$eatout,printAll = TRUE)
```

Fisher's Exact Test is conducted to supplement the results of the Chi-squared test from before. Based on this test, which compared two groups at a time, we can see that the there is a difference in terms of the frequency of eating out between white and black people as shown by the the low p-value of about 0.0003864 . Likewise, the frequency of eating out between white people and others is also difference based on the low p-value of 0.3543 . However, this test also shows that there is no difference between the frequency of eating out of black people and others given that the p-value is 
0.2815.

### Discussion and Conclusion


The question being responded to in this activity is whether there is a difference in the recurrence of eating out of different racial groups. The frist l test led is plotting a barplot of the information, and the assessed values show that there is a difference in the recurrence of eating out among the groups. This is additionally affirmed by the numerical results which showed the actual values in terms of percentages of people who eat out and do not eat out for each group. 
In conclusion, the Chi-squared test and the Fisher's Exact test affirmed that there is a difference between the recurrence of eating out among white and black people and others.In any case, it likewise guarantees that there isn't a difference in the data for black people and others. Generally speaking, we can reject the null hypothesis and claim that there is a difference in the frequency of eating out of various racial groups.







